---
title: "Thesis Final Markdown"
author: "Louisa Ong"
output:
  html_document:
    df_print: paged
---

# Setup  
  
```{r cache = TRUE}
# needed for fread
library(data.table)
library(dplyr)

nypd  <- fread("data/NYPD/nypd_brooklyn.csv", header = TRUE)

# functions located in another R sheet
source("thesis-functions.R")

```

    
# Datasets Being Used 

```{r cache = TRUE, message = FALSE, warning = FALSE}

# STEP 1: getting data for each year per tract - crime (nypd) and gentrification variables (census)
# 1.1 - different types of felonies
bk_count_2010 <- crimeCount(2010) # crime 
bk_count_2011 <- crimeCount(2011) # crime 
bk_count_2012 <- crimeCount(2012) # crime 
bk_count_2013 <- crimeCount(2013) # crime 
bk_count_2014 <- crimeCount(2014) # crime 
bk_count_2015 <- crimeCount(2015) # crime 
bk_count_2016 <- crimeCount(2016) # crime 
bk_count_2017 <- crimeCount(2017) # crime 
bk_count_2018 <- crimeCount(2018) # crime 
# 1.2 - getting census variables for calculating gentrification score
bk_gentri_2010 <- bkGentri(2010) # gentrification 
bk_gentri_2011 <- bkGentri(2011) # gentrification
bk_gentri_2012 <- bkGentri(2012) # gentrification
bk_gentri_2013 <- bkGentri(2013) # gentrification
bk_gentri_2014 <- bkGentri(2014) # gentrification
bk_gentri_2015 <- bkGentri(2015) # gentrification
bk_gentri_2016 <- bkGentri(2016) # gentrification
bk_gentri_2017 <- bkGentri(2017) # gentrification
bk_gentri_2018 <- bkGentri(2018) # gentrification

# STEP 2: Calculate the gentrification score 
# changes between variables between two years by the 750 tracts
# gentrification change rates for each year 2010-2018
bk_gen_2010_2011 <- calcChange(bk_gentri_2010, bk_gentri_2011)
bk_gen_2011_2012 <- calcChange(bk_gentri_2011, bk_gentri_2012)
bk_gen_2012_2013 <- calcChange(bk_gentri_2012, bk_gentri_2013)
bk_gen_2013_2014 <- calcChange(bk_gentri_2013, bk_gentri_2014)
bk_gen_2014_2015 <- calcChange(bk_gentri_2014, bk_gentri_2015)
bk_gen_2015_2016 <- calcChange(bk_gentri_2015, bk_gentri_2016)
bk_gen_2016_2017 <- calcChange(bk_gentri_2016, bk_gentri_2017)
bk_gen_2017_2018 <- calcChange(bk_gentri_2017, bk_gentri_2018)
bk_gen_all <- calcChange(bk_gentri_2010, bk_gentri_2018)

# STEP 3: Combine gentrification and crime 
# merging tracts 
# variables by the 750 tracts AND neighborhoods between two years
bk_gen_crime_2010 <- mergeTracts(bk_gen_2010_2011, bk_count_2010, bk_count_2011)
bk_gen_crime_2011 <- mergeTracts(bk_gen_2011_2012, bk_count_2011, bk_count_2012)
bk_gen_crime_2012 <- mergeTracts(bk_gen_2012_2013, bk_count_2012, bk_count_2013)
bk_gen_crime_2013 <- mergeTracts(bk_gen_2013_2014, bk_count_2013, bk_count_2014)
bk_gen_crime_2014 <- mergeTracts(bk_gen_2014_2015, bk_count_2014, bk_count_2015)
bk_gen_crime_2015 <- mergeTracts(bk_gen_2015_2016, bk_count_2015, bk_count_2016)
bk_gen_crime_2016 <- mergeTracts(bk_gen_2016_2017, bk_count_2016, bk_count_2017)
bk_gen_crime_2017 <- mergeTracts(bk_gen_2017_2018, bk_count_2017, bk_count_2018)
bk_gen_crime_all <- mergeTracts(bk_gen_all, bk_count_2010, bk_count_2018)

# STEP 4: Keep Neighborhood tracts as granularity
# Calculate crime count and gentrification scores for neighborhood level 
# 50 Neighborhoods, Gentrification Score, Each year (2010 is the earliest) for 2010-2018
bkg_2010 <- calcScore(bk_gen_crime_2010)
bkg_2011 <- calcScore(bk_gen_crime_2011)
bkg_2012 <- calcScore(bk_gen_crime_2012)
bkg_2013 <- calcScore(bk_gen_crime_2013)
bkg_2014 <- calcScore(bk_gen_crime_2014)
bkg_2015 <- calcScore(bk_gen_crime_2015)
bkg_2016 <- calcScore(bk_gen_crime_2016)
bkg_2017 <- calcScore(bk_gen_crime_2017)
bkg_all <- calcScore(bk_gen_crime_all)


# for felonies and census tract level 
bk_count_ofn_10 <- crimeCount_2(2010)
bk_count_ofn_18 <- crimeCount_2(2018)
bk_gen_crime_ofn_all <- mergeTracts_2(bk_gen_all, bk_count_ofn_10, bk_count_ofn_18)
bkg_ofn <- calcScore_OFN(bk_gen_crime_ofn_all)
bkg_tract <- calcScore_CT(bk_gen_crime_all)

```

# Keeping Non-Gentrified Areas  
  
```{r}
bkg_no_gentri <- bkg_all %>% 
  filter(NTA_GENTRI_SCORE < 0)

# without greenpoint and brooklyn cobble hill
bkg_no_gentri_2 <- bkg_no_gentri %>% 
  filter(! NTA_CODE %in% c("BK09", "BK76")) 

```


# Plot: Scatterplot of Gentrification & Crime - Neighborhood in Text  
  
```{r}
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - # 
# ------------- PLOT: Text Scatter of Gentri & Crime -------------- # 
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - # 

# distinguishing gentrified and less gentrified neighborhoods (binary)

# pts10 <- plotTextScatter(bkg_2010, "2010", "2011")
# pts11 <- plotTextScatter(bkg_2011, "2011", "2012")
# pts12 <- plotTextScatter(bkg_2012, "2012", "2013")
# pts13 <- plotTextScatter(bkg_2013, "2013", "2014")
# pts14 <- plotTextScatter(bkg_2014, "2014", "2015")
# pts15 <- plotTextScatter(bkg_2015, "2015", "2016")
# pts16 <- plotTextScatter(bkg_2016, "2016", "2017")
# pts17 <- plotTextScatter(bkg_2017, "2017", "2018")
pts_all <- plotTextScatter(bkg_all, "2010", "2018")
pts_no <- plotTextScatter(bkg_no_gentri, "2010", "2018")
pts_no_2 <- plotTextScatter(bkg_no_gentri_2, "2010", "2018") 


# ggplot() + 
#   geom_boxplot(data = bkg_all, aes(x = FEL_NTA_CHANGE)) + 
  


# pts_grid <- plot_grid(pts10, pts11, pts12,
#                       pts13, pts14, pts15,
#                       pts16, pts17, pts_all,
#                       ncol = 3)

# all gentri
pts_all + 
  labs(title = "\n \nText Scatterplot of All Neighborhoods", 
       subtitle = "Positioned by Gentrification Index Score and Change in Crime Rates (2010 - 2018)\n \n")  
  
# filtered
pts_filtered <- pts_no_2 + 
  xlim(min(-6), max(1)) + 
  geom_vline(xintercept = 0, color = "#333333") + 
  labs(title = "\nText Scatterplot of Non-Gentrified Neighborhoods", 
       subtitle = "Gentrification Index Score and Change in Crime Rates (2010 - 2018)\n")


# ggsave("visualizations/plot_text_scatter_grid.pdf", pts_grid, width = 12, height = 1)

# ggsave("visualizations/Text Scatter/all-gentri-text-scatter.svg", width = 8, height = 5)
# 
# ggsave("visualizations/Text Scatter/no-gentri-filt-text-scatter.svg", pts_filtered, width = 8, height = 5)


```


# Higher Granularity - Felony Types 

```{r}
# offense
bkg_ofn_df <- bkg_ofn %>% 
  as.data.frame()

fel_change_2 <- bkg_ofn_df %>% 
  filter(NTA_CODE %in% c("BK31", "BK34")) %>%
  group_by(FELONY_TYPE) %>%
  summarise(mean(OFN_NTA_CHANGE))
  
fel_change_all <- bkg_ofn_df %>% 
  group_by(FELONY_TYPE) %>%
  summarise(mean(OFN_NTA_CHANGE))


ftypes <- ggplot() +
  geom_point(data = fel_change_all, aes(x = FELONY_TYPE, y = `mean(OFN_NTA_CHANGE)`, fill = "All Boroughs"), color = "black") + 
  geom_point(data = fel_change_2, aes(x = FELONY_TYPE, y = `mean(OFN_NTA_CHANGE)`, fill = "Bay Ridge & Sunset Park"), color = "cyan4", alpha = 0.5) +
  # scale_fill_manual(values = c("darksalmon", "palegreen3", "honeydew3"), name = "Quadrant\nSignificance") +
  geom_hline(yintercept = 0, color = "grey70") + 
  coord_flip() +
  theme_minimal() +
  labs(title = "\nDifferent Types of Felonies among Boroughs",
      subtitle = "Change in Crime Rates (2010 - 2018)\n",
      y = "\nAverage Change in Crime Rate\n",
      x = "\nTypes of Felonies\n") +
  theme(text = element_text(family = "raleway", size = 12), 
        plot.title = element_text(family = "montserrat", size = 18))
  


ggsave("visualizations/felony-types.svg", ftypes, width = 12, height = 5)

```


# Dataframe of Binary Values - Gentrified vs. Not Gentrified, for each year    
  
Trend Lines   
  
```{r warning = FALSE}
# Making it Binary 
# using ntaMean function
nta_mean <- ntaMean(bkg_2011, 2011) %>% 
  bind_rows(ntaMean(bkg_2012, 2012)) %>%
  bind_rows(ntaMean(bkg_2013, 2013)) %>%
  bind_rows(ntaMean(bkg_2014, 2014)) %>% 
  bind_rows(ntaMean(bkg_2015, 2015)) %>% 
  bind_rows(ntaMean(bkg_2014, 2016)) %>% 
  bind_rows(ntaMean(bkg_2015, 2017))

# 2018 
# ser up empty dataset
nta_mean_2 <- data.frame(gentrified = "FALSE", 
                         avg_gentri = 0, 
                         avg_crime_rate_1 = nta_mean$avg_crime_rate_2[[13]], 
                         avg_crime_rate_2 = 0,
                         crime_change = 0,
                         # avg_ofn_rate_1 = 0,
                         # avg_ofn_rate_2 = 0,
                         # ofn_change = 0,
                         year = 2018) 

nta_mean_3 <- data.frame(gentrified = "TRUE", 
                         avg_gentri = 0, 
                         avg_crime_rate_1 = nta_mean$avg_crime_rate_2[[14]], 
                         avg_crime_rate_2 = 0, 
                         crime_change = 0,
                         # avg_ofn_rate_1 = 0,
                         # avg_ofn_rate_2 = 0,
                         # ofn_change = 0,
                         year = 2018) 

nta_mean_final <- nta_mean %>% 
  rbind(nta_mean_2) %>% 
  rbind(nta_mean_3)

# write.csv(nta_mean_final, "nta-gentri-crime-2010-2018.csv")
beep()
```


# Time Series Visualization  
  
```{r message = FALSE, warning = FALSE}

g_trend <- ggplot() +
  # plot points
  geom_point(data = nta_mean, 
             x = nta_mean$year,
             y = nta_mean$avg_crime_rate_1*100,
             group = nta_mean$gentrified, 
             color = "grey60") +
  # plot trend line connecting points
  geom_line(aes(x = nta_mean$year,
                y = nta_mean$avg_crime_rate_1*100,
                group = nta_mean$gentrified,
                color = nta_mean$gentrified),
                size = 1.02) +
  # legend label 
  scale_color_discrete(name = "Gentrification", labels = c("Absent", "Present")) +
  # tbeme elemnts
  theme_minimal() +
  theme(text = element_text(family = "raleway"),
        plot.title = element_text(family = "montserrat"))


g_trend_curve <- g_trend + 
  stat_smooth(aes(x = nta_mean$year,
                  y = nta_mean$avg_crime_rate_1*100, 
                   group = nta_mean$gentrified), 
              method = "lm", 
              formula = y ~ x + I(x^2), 
              size = 1, 
              color = "white", 
              alpha = 0.3) + 
  scale_x_continuous(breaks = nta_mean$year) +
  labs(title = "\nCrime in Brooklyn between\nGentrified & Non-Gentrified Neighborhoods (2010 - 2018)\n", 
       x = "\nYear\n",
       y = "\nAverage Crime Rate (Felony)\n") 
# ggsave("visualizations/trend_analysis_curve.svg", g_trend_curve, width = 12, height = 9)
beep()
```



# Time Series all trend lines

```{r}

nta_mean <- ntaMean(bkg_2011, 2011) %>% 
  bind_rows(ntaMean(bkg_2012, 2012)) %>%
  bind_rows(ntaMean(bkg_2013, 2013)) %>%
  bind_rows(ntaMean(bkg_2014, 2014)) %>% 
  bind_rows(ntaMean(bkg_2015, 2015)) %>% 
  bind_rows(ntaMean(bkg_2014, 2016)) %>% 
  bind_rows(ntaMean(bkg_2015, 2017))


bkg_2010$YEAR <- "2010"
bkg_2011$YEAR <- "2011"
bkg_2012$YEAR <- "2012"
bkg_2013$YEAR <- "2013"
bkg_2014$YEAR <- "2014"
bkg_2015$YEAR <- "2015"
bkg_2016$YEAR <- "2016"
bkg_2017$YEAR <- "2017"

bkg_all_no_gentri <- bkg_all %>% 
  filter(NTA_GENTRI_SCORE < 0)
bkg_all_no_gentri$YEAR <- NA
bkg_all_no_gentri$GENTRI <- FALSE


bkg_all_gentri <- bkg_all %>% 
  filter(NTA_GENTRI_SCORE >= 0)
bkg_all_gentri$YEAR <- NA 
bkg_all_gentri$GENTRI <- TRUE


bkg_years <- bkg_2010 %>% 
  bind_rows(bkg_2011) %>%
  bind_rows(bkg_2012) %>%
  bind_rows(bkg_2013) %>%
  bind_rows(bkg_2014) %>%
  bind_rows(bkg_2015) %>%
  bind_rows(bkg_2016) %>%
  bind_rows(bkg_2017)

# write.csv(bkg_years, "data/bkg_years.csv")

bkg_years_no_gentri <- bkg_years %>% 
  semi_join(bkg_all_no_gentri, "NTA_CODE") %>% 
  drop_na("YEAR") # needs tidyverse

# bkg_years_nta <- bkg_years %>% 
#   dplyr::select(-OFN_NTA_PAX_1, -OFN_NTA_PAX_2, -OFN_NTA_CHANGE) %>% 
#   distinct()

top_10_gentri <- bkg_all_no_gentri %>% 
  as.data.frame() %>%
  filter(! NTA_NAME %in% c("Brooklyn Heights-Cobble Hill", "Greenpoint")) %>%
  arrange(desc(NTA_GENTRI_SCORE)) %>% 
  slice(1:10) 

top_10_crime <- bkg_all_no_gentri %>% 
  as.data.frame() %>%
  filter(! NTA_NAME %in% c("Brooklyn Heights-Cobble Hill", "Greenpoint")) %>%
  arrange(desc(FEL_NTA_CHANGE)) %>% 
  slice(1:10) 


low_5 <- bkg_all_no_gentri %>% 
  as.data.frame() %>%
  filter(! NTA_NAME %in% c("Brooklyn Heights-Cobble Hill", "Greenpoint")) %>%
  arrange(NTA_GENTRI_SCORE) %>% 
  slice(1:5) 


```

 
 
# Spatial Mapping Visualizations 

```{r cache = TRUE}
nta_map_2010 <- toSpatial_NTA(bkg_2010)
nta_map_2011 <- toSpatial_NTA(bkg_2011)
nta_map_2012 <- toSpatial_NTA(bkg_2012)
nta_map_2013 <- toSpatial_NTA(bkg_2013)
nta_map_2014 <- toSpatial_NTA(bkg_2014)
nta_map_2015 <- toSpatial_NTA(bkg_2015)
nta_map_2016 <- toSpatial_NTA(bkg_2016)
nta_map_2017 <- toSpatial_NTA(bkg_2017) 

nta_map_all <- toSpatial_NTA(bkg_all)
nta_map_no_gentri <- toSpatial_NTA(bkg_no_gentri)
nta_map_no_gentri_2 <- toSpatial_NTA(bkg_no_gentri_2)

nta_map_no_gentri_spa <- nta_map_no_gentri_2 %>% 
  as_Spatial()
# writeOGR(obj = nta_map_no_gentri_spa, dsn = "nta_map_no_gentri", layer = "nta_map_no_gentri", driver = "ESRI Shapefile")

bkg_years_no_gentri_spa <- bkg_years_no_gentri %>% 
  toSpatial_NTA() %>%
  as_Spatial()

# writeOGR(obj = bkg_years_no_gentri_spa, dsn = "nta_all_years", layer = "nta_all_years", driver = "ESRI Shapefile")


ct_map <- toSpatial_tracts(bkg_tract)
ct_map_no_gentri <- bkg_tract %>%
  semi_join(bkg_no_gentri_2, join_by = "NTA_CODE") %>%
  filter(! NTA_CODE %in% c("BK09", "BK76")) %>%
  toSpatial_tracts()
ct_map_no_gentri_spa <- ct_map_no_gentri %>%
  as_Spatial()

# writeOGR(obj = ct_map_no_gentri_spa, dsn = "ct_map_no_gentri", layer = "ct_map_no_gentri", driver = "ESRI Shapefile")

beep()

nta_poly <- st_read("nta_map_no_gentri/nta_map_no_gentri.shp")
ct_poly <- st_read("ct_map_no_gentri/ct_map_no_gentri.shp")
all_years_poly <- st_read("nta_all_years/nta_all_years.shp")

names(nta_poly)
```
  
  
#  Boxplot

```{r}

bp_ct_all <- ggplot(data = ct_map) + 
  geom_boxplot(aes(x = FEL_CT_CHANGE)) + 
  ylim(-3, 3) + 
  labs(title = "\nBoxplot of Change in Crime Rates for All Tracts",
      subtitle = "(2010 - 2018)\n",
      x = "\nChange in Crime Rates by Tract\n") +
  theme_minimal() +
  geom_vline(xintercept = 0, color = "grey50") +
  theme(text = element_text(family = "raleway", size = 14), 
        plot.title = element_text(family = "montserrat", size = 17))



bp_ct <- ggplot(data = ct_map_no_gentri) + 
  geom_boxplot(aes(x = FEL_CT_CHANGE)) + 
  ylim(-3, 3) + 
  labs(title = "\nBoxplot of Change in Crime Rates for Non-Gentrified Tracts",
      subtitle = "(2010 - 2018)\n",
      x = "\nChange in Crime Rates by Tract\n") +
  theme_minimal() +
  geom_vline(xintercept = 0, color = "grey50") +
  theme(text = element_text(family = "raleway", size = 14), 
        plot.title = element_text(family = "montserrat", size = 17))
        # plot.margin=unit(c(3,3,3,3),"cm"))

bp_nta <- ggplot(data = nta_map_no_gentri_2) + 
  geom_boxplot(aes(x = FEL_NTA_CHANGE)) + 
  labs(title = "\nBoxplot of Change in Crime Rates for Non-Gentrified Neighborhoods",
    subtitle = "(2010 - 2018)\n",
    x = "\nChange in Crime Rates by Neighborhood\n") + 
  ylim(-3, 3) + 
  theme_minimal() +
  geom_vline(xintercept = 0, color = "grey50") +
  theme(text = element_text(family = "raleway", size = 14), 
      plot.title = element_text(family = "montserrat", size = 17)) 


bp_all <- ggplot(data = nta_map_all) + 
  geom_boxplot(aes(x = FEL_NTA_CHANGE)) + 
  labs(title = "\nBoxplot of Change in Crime Rates for All Neighborhoods",
    subtitle = "(2010 - 2018)\n",
    x = "\nChange in Crime Rates by Neighborhood\n") + 
  ylim(-3, 3) + 
  xlim(-0.004, 0.0025) +
  theme_minimal() +
  geom_vline(xintercept = 0, color = "grey50") +
  theme(text = element_text(family = "raleway", size = 14), 
      plot.title = element_text(family = "montserrat", size = 17)) 

      # plot.margin=unit(c(3,3,3,3),"cm"))

# bp_grid <- plot_grid(bp_ct, bp_nta, n_col = 0)

# ggplot(data = ct_map_no_gentri) + 
#   geom_sf(aes(fill = FEL_CT_CHANGE))

# ggsave("visualizations/box_plot_fel_both.svg", bp_grid, width = 8, height = 15)
# ggsave("visualizations/box_plot_fel_tract.svg", bp_ct, width = 10, height = 4)
# ggsave("visualizations/box_plot_fel_tract_all.svg", bp_ct_all, width = 10, height = 4)
  # ggsave("visualizations/box_plot_fel_nta.svg", bp_nta, width = 10, height = 4)
# ggsave("visualizations/box_plot_fel_all.svg", bp_all, width = 10, height = 4)

```

  
# Regression Models

```{r}

fit_ct_basic <- lm(FEL_CT_C ~ CT_GE, data = ct_poly) # No high school, No family, White 
fit_ct_multi <- lm(FEL_CT_C ~ CT_NO_H + CT_WH + CT_NO_F, data = ct_poly) # No high school, No family, White 

fit_nta_basic <- lm(FEL_NTA_C ~ NTA_G, data = nta_poly) # No high school, No family, White 
fit_nta_multi <- lm(FEL_NTA_C ~ NTA_NO_H + NTA_W + NTA_NO_F, data = nta_poly) # No high school, No family, White 
fit_nta_time <- plm(FEL_NTA_C ~ NTA_G + as.factor(YEAR), index = c("NTA_C", "YEAR"), data = all_years_poly)
                 
# stargazer(fit_nta, type = "text", digits = 5, out = "nta-fit-ols.txt")

fit <- stan_glm(FEL_CT_CHANGE ~ CT_GENTRI_SCORE, data = ct_map_no_gentri, refresh = 0) 

print(fit, digits = 6)

plot(fit_ct, which = 3)

resids <- residuals(fit_ct)
# resids <- residuals(ct_reg)

ct_reg <- as_Spatial(ct_map_no_gentri)

colours <- c("dark blue", "blue", "red", "dark red") 
#here it is assumed that your eastings and northings coordinates are stored in columns called x and y in your dataframe
map.resids <- SpatialPointsDataFrame(data=data.frame(resids), coords=cbind(x,y)) 

#for speed we are just going to use the quick sp plot function, but you could alternatively store your residuals back in your LondonWards dataframe and plot using geom_point in ggplot2
spplot(map.resids, cuts=quantile(resids), col.regions=colours, cex=1) 

```



```{r message = FALSE, warning = FALSE} 


ntm10 <- plotMapGentri(nta_map_2010, "2010", "2011")
ntm11 <- plotMapGentri(nta_map_2011, "2011", "2012")
ntm12 <- plotMapGentri(nta_map_2012, "2012", "2013")
ntm13 <- plotMapGentri(nta_map_2013, "2013", "2014")
ntm14 <- plotMapGentri(nta_map_2014, "2014", "2015")
ntm15 <- plotMapGentri(nta_map_2015, "2015", "2016")
ntm16 <- plotMapGentri(nta_map_2016, "2016", "2017")
ntm17 <- plotMapGentri(nta_map_2017, "2017", "2018")


ntm_all <- plotMapGentri(nta_map_all, "2010", "2018")
ntm_no_gentri <- plotMapGentri(nta_map_no_gentri_2, "2010", "2018")


ntm_grid <- plot_grid(ntm10, ntm11, ntm12,
                      ntm13, ntm14, ntm15,
                      ntm16, ntm17, ntm_all,
                      ncol = 3)
# ggsave("visualizations/plot_map_nta_gentri.pdf", ntm_grid, width = 12, height = 10)
# ggsave("visualizations/map_all_gentri_fel.pdf", width = 12, height = 7)
ntm_no_gentri_2 <- ntm_no_gentri + 
  geom_sf(data = nta_map_all, aes(color = "honeydew3"), color = NA) +
  geom_sf(data = nta_map_no_gentri, aes(fill = nta_map_no_gentri$NTA_GENTRI_SCORE), color = NA)



beep()
pmc10 <- plotMapCrime(nta_map_2010, "2010", "2011")
pmc11 <- plotMapCrime(nta_map_2011, "2011", "2012")
pmc12 <- plotMapCrime(nta_map_2012, "2012", "2013")
pmc13 <- plotMapCrime(nta_map_2013, "2013", "2014")
pmc14 <- plotMapCrime(nta_map_2014, "2014", "2015")
pmc15 <- plotMapCrime(nta_map_2015, "2015", "2016")
pmc16 <- plotMapCrime(nta_map_2016, "2016", "2017")
pmc17 <- plotMapCrime(nta_map_2017, "2017", "2018")
pmc_all <- plotMapCrime(nta_map_all, "2010", "2018")
pmc_no_gentri <- plotMapCrime(nta_map_no_gentri_2, "2010", "2018")
                        
pmc_grid <- plot_grid(pmc10, pmc11, pmc12,
                      pmc13, pmc14, pmc15,
                      pmc16, pmc17, pmc_all,
                      ncol = 3)
# ggsave("visualizations/plot_map_nta_crime.pdf", pmc_grid, width = 12, height = 10)
nta_map_no_gentri <- nta_map_no_gentri %>% 
  filter(! NTA_CODE %in% c("BK09", "BK76"))


ntm_all <- ntm_all + 
  labs(title = "Gentrification in Brooklyn",
       subtitle = "2010 - 2018\n") + 
  theme(text = element_text(family = "raleway", size = 12), 
        plot.title = element_text(family = "montserrat", size = 18))  


pmc_all <- pmc_all + 
  labs(title = "Change in Crime Rates in Brooklyn",
       subtitle = "2010 - 2018\n") + 
  theme(text = element_text(family = "raleway", size = 12), 
        plot.title = element_text(family = "montserrat", size = 18))


pmc_no_gentri_both_fel <- pmc_no_gentri + 
  geom_sf(data = nta_map_all, aes(color = "honeydew3"), color = NA) + 
  geom_sf(data = nta_map_no_gentri_2, aes(fill = nta_map_no_gentri_2$FEL_NTA_CHANGE), color = NA) + 
  scale_fill_viridis_c(option = "magma", name = "Change in Crime Rates", direction = -1) + 
  labs(title = "Change in Crime Rates for\nNon-Gentrified Neighborhoods",
      subtitle = "(2010 - 2018)\n") + 
  theme(text = element_text(family = "raleway", size = 14), 
        plot.title = element_text(family = "montserrat", size = 17))
        # plot.margin=unit(c(1,1,1.5,1.2),"cm"))


pmc_no_gentri_both_gentri <- pmc_no_gentri + 
  geom_sf(data = nta_map_all, aes(color = "honeydew3"), color = NA) + 
  geom_sf(data = nta_map_no_gentri_2, aes(fill = nta_map_no_gentri_2$NTA_GENTRI_SCORE), color = NA) +
  scale_fill_viridis_c(option = "magma", name = "Gentrification Score", direction = 1) +
  labs(title = "Gentrification Score for\nNon-Gentrified Neighborhoods",
       subtitle = "(2010 - 2018)\n") +
  theme(text = element_text(family = "raleway", size = 14),
        plot.title = element_text(family = "montserrat", size = 17))


all_grid_both <- plot_grid(ntm_all, pmc_all, ncol = 2)

both_grid <- plot_grid(pmc_no_gentri_both_gentri, pmc_no_gentri_both_fel, ncol = 2)
# 
ggsave("visualizations/Map/map_all_gentri.svg", ntm_all, width = 8, height = 5)
ggsave("visualizations/Map/map_all_crime.svg", pmc_all, width = 8, height = 5)
ggsave("visualizations/Map/all_grid_both.svg", all_grid_both, width = 9, height = 5)

ggsave("visualizations/Map/map_grid_gentri.svg", ntm_grid, width = 14, height = 11)
ggsave("visualizations/Map/map_grid_fel.svg", pmc_grid, width = 14, height = 11)
ggsave("visualizations/Map/map_both.svg", both_grid, width = 9, height = 5)

ggsave("visualizations/Map/map_no_gentri_crime.svg", pmc_no_gentri_both, width = 8, height = 5)
# 
# 

```


# Spatial Autocorrelation

```{r}

map_analysis <- nta_map_no_gentri_2 %>%
  as_Spatial()

# map_analysis <- nta_map_all %>%
  # as_Spatial()

# Queens Contiguity Matrix
spat_matrix <- poly2nb(map_analysis)
# Neighbors list with Spatial Weihgts
list_w <- nb2listw(spat_matrix)
# calculate local Moran of the distribution 
lmoran_f <- localmoran(map_analysis$FEL_NTA_CHANGE, list_w)
lmoran_g <- localmoran(map_analysis$NTA_GENTRI_SCORE, list_w)

sum_f <- summary(lmoran_f)

summary(lmoran_g)
stargazer(lmoran_f, summary = TRUE, out = "lmoran_f.png")


# padronize the variable and save it to a new column 
map_analysis$s_fel <- scale(map_analysis$FEL_NTA_CHANGE) %>% as.vector()
map_analysis$s_gentri <- scale(map_analysis$NTA_GENTRI_SCORE) %>% as.vector()
# create a spatially lagged variable and save it to a new column 
map_analysis$lag_s_fel <- lag.listw(list_w, map_analysis$s_fel)
map_analysis$lag_s_gentri <- lag.listw(list_w, map_analysis$s_gentri)
# summary of variables to form the analysis 
summary(map_analysis$s_fel)
summary(map_analysis$lag_s_fel)
summary(map_analysis$s_gentri)
summary(map_analysis$lag_s_gentri)


# moran scatterplot, in basic graphics - with identificaiton of influential observations 
x_fel <- map_analysis$s_fel
y_fel <- map_analysis$lag_s_fel %>% as.vector()
xx_fel <- data.frame(x_fel, y_fel)
x_gentri <- map_analysis$s_gentri
y_gentri <- map_analysis$lag_s_gentri %>% as.vector()
xx_gentri <- data.frame(x_gentri, y_gentri)
moran.plot(x_fel, list_w)
moran.plot(x_gentri, list_w)

xx <- data.frame(x = x_fel, y = y_fel)
ggplot(xx, aes(x_fel, y_fel)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = F, color = "hotpink4") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "deepskyblue4") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "deepskyblue4") + 
  theme_minimal() + 
  labs(title = "\nMoran's I Scatterplot",
       subtitle = "Felony for Non-Gentrified Neighborhoods\n",
       x = "\nFelony\n", 
       y = "\nSpatially Lagged Felony\n") + 
  theme(text = element_text(family = "raleway", size = 12), 
        plot.title = element_text(family = "montserrat", size = 18, color = "grey20"))
  

# QUADRANTS 
map_analysis$quad_sig_fel <- NA
map_analysis$quad_sig_gentri <- NA
# high high quadrant 
map_analysis[(map_analysis$s_fel >= 0 & 
              map_analysis$lag_s_fel >= 0) & 
               (lmoran_f[, 5] <= 0.05), "quad_sig_fel"] <- "high-high"
# low low  quadrant 
map_analysis[(map_analysis$s_fel <= 0 & 
              map_analysis$lag_s_fel <= 0) & 
               (lmoran_f[, 5] <= 0.05), "quad_sig_fel"] <- "low-low"
# high low  quadrant 
map_analysis[(map_analysis$s_fel >= 0 & 
              map_analysis$lag_s_fel <= 0) & 
               (lmoran_f[, 5] <= 0.05), "quad_sig_fel"] <- "highlow"
# low low  quadrant 
map_analysis[(map_analysis$s_fel <= 0 & 
              map_analysis$lag_s_fel >= 0) & 
               (lmoran_f[, 5] <= 0.05), "quad_sig_fel"] <- "low-high"
# FOR GENTRI
# high high quadrant 
map_analysis[(map_analysis$s_gentri >= 0) & 
               (lmoran_g[, 5] <= 0.05), "quad_sig_gentri"] <- "high-high"

# low low  quadrant 
map_analysis[(map_analysis$s_gentri <= 0 & 
              map_analysis$lag_s_gentri <= 0) & 
               (lmoran_g[, 5] <= 0.05), "quad_sig_gentri"] <- "low-low"

# high low  quadrant 
map_analysis[(map_analysis$s_gentri >= 0 & 
              map_analysis$lag_s_gentri <= 0) & 
               (lmoran_g[, 5] <= 0.05), "quad_sig_gentri"] <- "highlow"
# low low  quadrant 
map_analysis[(map_analysis$s_gentri <= 0 & 
              map_analysis$lag_s_gentri >= 0) & 
               (lmoran_g[, 5] <= 0.05), "quad_sig_gentri"] <- "low-high"
# non-significant observations 
# map_analysis <- as_Spatial(map_analysis)
map_analysis@data[(lmoran_f[, 5] > 0.05), "quad_sig_fel"] <- "not signif."
map_analysis@data[(lmoran_g[, 5] > 0.05), "quad_sig_gentri"] <- "not signif."
map_analysis$quad_sig_fel <- as.factor(map_analysis$quad_sig_fel)
map_analysis$quad_sig_gentri <- as.factor(map_analysis$quad_sig_gentri)
map_analysis@data$id <- rownames(map_analysis@data)
# plotting the map
df <- fortify(map_analysis, region = "id")
df <- left_join(df, map_analysis@data)

lisa_fel_plot <- df %>% 
  ggplot(aes(long, lat, group = group, fill = quad_sig_fel)) + 
  geom_polygon(color = "white", size = .05) + 
  coord_equal() +
  labs(title = "\nLISA Clusters for\nChange in Felony Crime Rates (2010-2018)") + 
  scale_fill_manual(values = c("darksalmon", "palegreen3", "honeydew3"), name = "Quadrant\nSignificance") +
  theme_map() + 
  theme(text = element_text(family = "raleway", size = 12), 
        plot.title = element_text(family = "montserrat", size = 18, color = "grey20"))

lisa_gentri_plot <- df %>% 
  ggplot(aes(long, lat, group = group, fill = quad_sig_gentri)) + 
  geom_polygon(color = "white", size = .05) + 
  coord_equal() +
  labs(title = "\nLISA Clusters for\nChange in Gentrification Score (2010-2018)") + 
  scale_fill_manual(values = c("honeydew3"), name = "Quadrant\nSignificance") +
  theme_map() + 
  theme(text = element_text(family = "raleway", size = 12), 
        plot.title = element_text(family = "montserrat", size = 18, color = "grey20"))

lisa_grid <- plot_grid(lisa_fel_plot, lisa_gentri_plot,
                      ncol = 2)

# ggsave("visualizations/lisa-gentri-no-gentri.svg", lisa_gentri_plot, width = 8, height = 5)
# 
# ggsave("visualizations/morans-i.svg", width = 8, height = 5)
# 
# ggsave("visualizations/lisa-fel-no-gentri-2.svg", lisa_fel_plot, width = 8, height = 5)
# 
# ggsave("visualizations/lisa-grid-no-gentri.svg", lisa_grid, width = 13, height = 8)
# ```


